---
output: github_document
params:
  maintainer: Emma Mendelsohn
  maintainer_email: schwantes@ecohealthalliance.org
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# wahisdb

[![License (for code): MIT](https://img.shields.io/badge/License (for code)-MIT-green.svg)](https://opensource.org/licenses/MIT)
[![License: CC0-1.0](https://img.shields.io/badge/License (for data)-CC0_1.0-lightgrey.svg)](http://creativecommons.org/publicdomain/zero/1.0/)

This package accesses and formats veterinary disease data from [OIE WAHIS](https://wahis.woah.org/#/home). Data is updated weekly and is publicly available on DoltHub: https://www.dolthub.com/repositories/ecohealthalliance/wahisdb.

## Database Tables
- **outbreak_reports_ingest_status_log** List of ingested reports. If report ingestion failed, the error will be in the `ingest_error` field. `report_info_id` is the unique primary key for this table. It represents the base values of the API URL for the report (see the `url` field). `dashboard_report_id` is the report ID as it appears in the WAHIS dashboard, which does NOT match report_id the tables retrieved from the API.
- **outbreak_reports_events_raw** Raw (uncleaned/standardized) data from the WAHIS API. Each row is an outbreak report, with high level information on the country, disease, and disease status. `report_id` is the unique primary key. Note that this is an API-specific report ID and does NOT match the report ID shown on the WAHIS dashboard. This table can be joined with outbreak_reports_ingest_status_log by `report_info_id`.
- **outbreak_reports_details_raw** Raw (uncleaned/standardized) data from the WAHIS API. Detailed location and impact data for outbreak subevents (e.g., individual farms within a larger outbreak event). `unique_id` is a generated unique primary key, consisting of a concatenation of report_id, oie_reference (i.e., outbreak_location_id), and species_name (i.e., taxon). This table can be joined with outbreak_reports_events_raw by `report_id`
- **outbreak_summary** Summarizes outbreak data by event/thread, with cleaned and standardized fields. Each row in an outbreak event. `outbreak_thread_id` is the unique primary key, which matches `event_id_oie_reference` in outbreak_reports_ingest_status_log.
- **outbreak_time_series** a A cleaned/standardized version of outbreak_reports_details_raw. Detailed location and impact data for outbreak subevents (e.g., individual farms within a larger outbreak event). `unique_id` is a generated unique primary key, consisting of a concatenation of report_id, outbreak_location_id, and  taxon.   `unique_id`  matches one to one with outbreak_reports_details_raw. This table can be joined with outbreak_summary by `outbreak_thread_id'.
- **disease_key** Hand-curated lookup for disease name standardization and taxonomy, used to clean the disease names in outbreak_summary. `disease` is the primary key, and can be used to join with outbreak summary. 

## Repository Structure and Reproducibility

-   `wahisdb/` contains the dolt database. See instructions below. 
-   `R/` contains functions used in this analysis. 
-   This project uses the `targets` package to create its analysis pipeline.  The steps are
    defined in the `_targets.R` file and the workflow can be executed by running `targets::tar_make()`.
-   The schematic figure below summarizes the steps. (The figure is generated using
    `mermaid.js` syntax and should display as a graph on GitHub.It can also be viewed by        pasting the code into <https://mermaid.live>.)

```{r, echo=FALSE, message = FALSE, results='asis'}
mer <- targets::tar_mermaid(targets_only = TRUE, outdated = FALSE, 
                            legend = FALSE, color = FALSE, 
                            exclude = c("readme", ends_with("_targets")))
cat(
  "```mermaid",
  mer[1], 
  #'Objects([""Objects""]) --- Functions>""Functions""]',
  'subgraph Project Workflow',
  mer[3:length(mer)],
  'linkStyle 0 stroke-width:0px;',
  "```",
  sep = "\n"
)
```


## Dolt
-   Install and configure the database software, dolt: https://www.dolthub.com/blog/2020-02-03-dolt-and-dolthub-getting-started/
-   Install: sudo curl -L https://github.com/dolthub/dolt/releases/latest/download/install.sh | sudo bash
-   Provide credentials: dolt config --global --add user.email YOU@DOMAIN.COM and dolt config --global --add user.name "YOUR NAME"
Login: dolt login
-   Copy key to https://www.dolthub.com/settings/credentials
-   Clone the dolt database: dolt clone ecohealthalliance/wahisdb

## renv
-   This project uses the [{renv}](https://rstudio.github.io/renv/) framework to 
    record R package dependencies and versions. Packages and versions used are recorded in 
    `renv.lock` and code used to manage dependencies is in `renv/` and other files
    in the root project directory.  On starting an R session in the working directory, run `renv::restore()` to install R package dependencies.
