---
output: github_document
params:
  maintainer: Emma Mendelsohn
  maintainer_email: schwantes@ecohealthalliance.org
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# wahisdb

[![License (for code): MIT](https://img.shields.io/badge/License (for code)-MIT-green.svg)](https://opensource.org/licenses/MIT)
[![License: CC0-1.0](https://img.shields.io/badge/License (for data)-CC0_1.0-lightgrey.svg)](http://creativecommons.org/publicdomain/zero/1.0/)

This package accesses and formats veterinary disease data from [OIE WAHIS](https://wahis.woah.org/#/home). Data is updated weekly and is publicly available on DoltHub: https://www.dolthub.com/repositories/ecohealthalliance/wahisdb.

## Database Tables
-   __outbreak_reports_ingest_status_log__ List of reports in database. `report_info_id` can be appended to "https://wahis.oie.int/pi/getReport/" to see the report API, and to "https://wahis.oie.int/#/report-info?reportId=" to see the formatted outbreak report.
-   __disease_key__ lookup for disease name standardization and taxonomy. Currently not being used within this workflow. 
-   __outbreak_reports_diseases_unmatched__ disease names that were not found in the ANDO disease name lookup.
-   __outbreak_reports_events_raw__ High-level event information including country, disease and disease status. Disease names are standardized to the [Animal Disease Ontology](http://agroportal.lirmm.fr/ontologies/ANDO]) from the French National research institute for agriculture, food and the environment. Each row is an outbreak report. `report_id` is the unique report ID.
-   __outbreak_summary__ a cleaned-up version of outbreak_reports_events_raw.
-   __outbreak_reports_details_raw__ Detailed location and impact data for outbreak events. This table can be joined with `outbreak_reports_events` by `report_id`. `outbreak_location_id` is a unique ID for each location (e.g, farm or village) within a outbreak. The field `id` is the unique combination of the `report_id`, `outbreak_location_id`, and `taxa`.
-   __outbreak_time_series__ a cleaned-up version of outbreak_reports_details_raw

## Repository Structure and Reproducibility

-   `wahisdb/` contains the dolt database. See instructions below. 
-   `R/` contains functions used in this analysis. 
-   This project uses the `targets` package to create its analysis pipeline.  The steps are
    defined in the `_targets.R` file and the workflow can be executed by running `targets::tar_make()`.
-   The schematic figure below summarizes the steps. (The figure is generated using
    `mermaid.js` syntax and should display as a graph on GitHub.It can also be viewed by        pasting the code into <https://mermaid.live>.)

```{r, echo=FALSE, message = FALSE, results='asis'}
mer <- targets::tar_mermaid(targets_only = TRUE, outdated = FALSE, 
                            legend = FALSE, color = FALSE, 
                            exclude = c("readme", ends_with("_targets")))
cat(
  "```mermaid",
  mer[1], 
  #'Objects([""Objects""]) --- Functions>""Functions""]',
  'subgraph Project Workflow',
  mer[3:length(mer)],
  'linkStyle 0 stroke-width:0px;',
  "```",
  sep = "\n"
)
```


## Dolt
-   Install and configure the database software, dolt: https://www.dolthub.com/blog/2020-02-03-dolt-and-dolthub-getting-started/
-   Install: sudo curl -L https://github.com/dolthub/dolt/releases/latest/download/install.sh | sudo bash
-   Provide credentials: dolt config --global --add user.email YOU@DOMAIN.COM and dolt config --global --add user.name "YOUR NAME"
Login: dolt login
-   Copy key to https://www.dolthub.com/settings/credentials
-   Clone the dolt database: dolt clone ecohealthalliance/wahisdb

## renv
-   This project uses the [{renv}](https://rstudio.github.io/renv/) framework to 
    record R package dependencies and versions. Packages and versions used are recorded in 
    `renv.lock` and code used to manage dependencies is in `renv/` and other files
    in the root project directory.  On starting an R session in the working directory, run `renv::restore()` to install R package dependencies.
